Investigating the applications and limits of single particle light scattering using limited detection schemes

### 2022-11-08 Manuscript.tex
Add files via upload

Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics~\cite{Wang 2021} to those that test the fundamental assumptions of Quantum Mechanics ~\cite{Duan 2013} thanks mainly to the ability of the tweezer to transduce and detect forces in the order of a few pico-newton. Recent applications of the Optical Tweezers in metrological measurements ~\cite{Scott Waitukaitis} and colloidal aggregation ~\cite{Optical Binding} demand the tweezer to also be able to characterise  the trapped entity. To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering ~\cite{P K Gupta} and dynamical characterisation has been achieved mainly by following the centre-of-mass Brownian motion of the trapped entity using a Quadrant Photo Detector~\cite{Alexander Rohrbach}. While this technique has allowed for detection of linear motion of trapped entities  with sub-nanometre precision~\cite{Alexander Rohrbach} and  rotation~\cite{recent paper} of the centre-of-mass, there is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in ~\cite{New Zealand paper} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.

Possibility of integrating light scattering with optical trapping was demonstrated by Bar-Ziv et.al.,~\cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's Brownian motion was studied. While this allowed for studying the dynamics, any information with regards structure of the trapped bead was precluded as the measurement was a single-angle scattering measurement.
In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in ~\cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure 1.

@@

We use the a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{MSTM} to determine the light scattering intensity of our dimer in each reference orientation. We assume that the for each detection fibre the intensities can be distributed on a Gaussian curve centred at $y_k$, the normalised intensity at detection fibre k. Based on the light scattering from the dimer we can assign a probability that the produced light scattering pattern produced belongs to the reference orientation $\hat{n}$

@@

\subsection{Testing the model}
\label{2.1}
We use the Brownian OT package by Jerome Fung \cite{Brownian_OT} to simulate how our target dimer moves while subject to an optical trap. At each time step we determine the dimer's orientation and use MSTM to calculate its scattering pattern. We then use the above method to estimate the dimer orientation based on its scattering pattern. The priori estimate $p(\hat{n})$ can be defined as a Boltzmann's distribution based on the normalised dot product of our previous estimation $\hat{n}_{t-\Delta t}$ and each reference orientation $\hat{n}_t$:

@@

\subparagraph{Optimization of parameters}
\label{2.2}
The efficacy of our estimation is a function of our detection angles, testing every possible configuration manually is time consuming, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model. Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Ultranest}. Ultra-nest implements a variation of Monte Carlo integration,nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).

@@

As can be seen from \figurename{ 2}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap. Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis. It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely. Further simulations of differently sized dimers showed similar results, but only when $a_1 \geq 2a_2$. Dimers with more symmetrical size ratios immediately aligned into a fixed vertical position. 
In Vigilante's work with dimers \cite{Brownian_OT} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations. Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.

>>>>

\indent Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics  \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces in the order of a few piconewton. Recent applications of the Optical Tweezers in metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical} demand the tweezer to also be able to characterise  the trapped entity. To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman} and dynamical characterisation has been achieved mainly by following the centre-of-mass brownian motion of the trapped entity using a Quadrant Photo Detector \cite{friedrich2012tuning}. While this technique has allowed for detection of linear motion of trapped entities  with sub-nanometre precision \cite{friedrich2012tuning} and  rotation \cite{yifat2021facile} of the centre-of-mass, there is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.

Possibility of integrating light scattering with optical trapping was demonstrated by Saffran and 
co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's brownian motion was studied. While this allowed for studying the dynamics, any information with regards structure of the trapped bead was precluded as the measurement was a single-angle scattering measurement.
In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure 1.

@@

We use the a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{Mishchenko1996MSTM} to determine the light scattering intensity of our dimer in each reference orientation. We assume that the for each detection fibre the intensities can be distributed on a Gaussian curve centred at $y_k$, the normalised intensity at detection fibre k. Based on the light scattering from the dimer we can assign a probability that the produced light scattering pattern produced belongs to the reference orientation $\hat{n}$

@@

\subsection{Testing the model}
\label{2.1}
We use the Brownian OT package by Jerome Fung \cite{Vigilante2020Brownian_OT} to simulate how our target dimer moves while subject to an optical trap. At each time step we determine the dimer's orientation and use MSTM to calculate its scattering pattern. We then use the above method to estimate the dimer orientation based on its scattering pattern. The priori estimate $p(\hat{n})$ can be defined as a Boltzmann's distribution based on the normalised dot product of our previous estimation $\hat{n}_{t-\Delta t}$ and each reference orientation $\hat{n}_t$:

@@

\subparagraph{Optimization of parameters}
\label{2.2}
The efficacy of our estimation is a function of our detection angles, testing every possible configuration manually is time consuming, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model. Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Buchner2016Ultranest}. Ultra-nest implements a variation of Monte Carlo integration,nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).

@@

As can be seen from \figurename{ 2}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap. Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis. It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely. Further simulations of differently sized dimers showed similar results, but only when $a_1 \geq 2a_2$. Dimers with more symmetrical size ratios immediately aligned into a fixed vertical position. 
In Vigilante's work with dimers \cite{Vigilante2020Brownian_OT} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations. Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.
